---
title: "ENV 790.30 - Time Series Analysis for Energy Data | Spring 2022"
subtitle: "Assignment 3 - Due date 02/08/22"
author: "Richard Zhang"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: inline
---
  
## Directions
  
You should open the .rmd file corresponding to this assignment on RStudio. The file is available on our class repository on Github. 

Once you have the project open the first thing you will do is change "Student Name" on line 3 with your name. 
Then you will start working through the assignment by  **creating code and output** that answer each question. Be sure to use this assignment document. Your report should contain the answer to each question and any plots/tables you obtained (when applicable).

Please keep this R code chunk options for the report. It is easier for us to grade when we can see code and output together. And the tidy.opts will make sure that line breaks on your code chunks are automatically added for better visualization. 

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

When you have completed the assignment, **Knit** the text and code into a single PDF file. Rename the pdf file such that it includes your first and last name (e.g., "LuanaLima_TSA_A03_Sp22.Rmd"). Submit this pdf using Sakai. 

## Questions 

Consider the same data you used for A2 from the spreadsheet "Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx". The data comes from the US Energy Information and Administration and corresponds to the January 2022 **Monthly** Energy Review. Once again you will work only with the following columns: Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption. Create a data frame structure with these three time series only. 

R packages needed for this assignment:"forecast","tseries", and "Kendall". Install these packages, if you haven't done yet. Do not forget to load them before running your script, since they are NOT default packages.\\

```{r}
#Load/install required package here
library(forecast)
library(tseries)
library(Kendall)
library(readxl)
library(ggplot2)
```

##Trend Component

### Q1
Create a plot window that has one row and three columns. And then for each object on your data frame, fill the plot window with time series plot, ACF and PACF. You may use the some code from A2, but I want all three plots on the same window this time. (Hint: use par() function)

```{r Import data}
energy_data <- read_excel("/Users/RichardZ/Downloads/ENV790_TimeSeriesAnalysis_Sp2022/Data/Table_10.1_Renewable_Energy_Production_and_Consumption_by_Source.xlsx", skip=10)

energy_data <- energy_data[-1,1:6]
energy_data <- energy_data[,-(2:3)]

nproduction <- ncol(energy_data)-1
nobservation <- nrow(energy_data) 

date <- as.Date(energy_data$Month, format = "%Y-%m-%d")

head(energy_data)
str(energy_data)
tail(energy_data)
```

```{r Q1}
ts_energy_data <- ts(energy_data[,2:4],start=date[1],frequency=12)

for(i in 1:nproduction){
  par(mfrow=c(1,3))
  plot(ts_energy_data[,i],type="l",col="black",ylab="Production [Trillion Btu]",main=colnames(energy_data)[i+1])
  Acf(ts_energy_data[,i],lag.max=40,main=paste(colnames(energy_data)[i+1],i,sep=""))
  Pacf(ts_energy_data[,i],lag.max=40,main=paste(colnames(energy_data)[i+1],i,sep=""))
}

```

### Q2
From the plot in Q1, do the series Total Biomass Energy Production, Total Renewable Energy Production, Hydroelectric Power Consumption appear to have a trend? If yes, what kind of trend?

```{r Q2}
t <- c(1:nobservation)
df_energy_data <- data.frame(energy_data)

for(i in 1:nproduction){
  df_energy_data[,i+1] <- as.numeric(df_energy_data[,i+1])
}
##Biomass
  linear_trend_model=lm(df_energy_data[,2]~t)  
  summary(linear_trend_model)
  Biobeta0=as.numeric(linear_trend_model$coefficients[1])  
  Biobeta1=as.numeric(linear_trend_model$coefficients[2])  

  ggplot(df_energy_data, aes(x=date, y=df_energy_data[,2])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(df_energy_data)[2],sep="")) +
            geom_smooth(color="red",method="lm")
##Renewable
  linear_trend_model=lm(df_energy_data[,3]~t)  
  summary(linear_trend_model)
  Renewbeta0=as.numeric(linear_trend_model$coefficients[1])  
  Renewbeta1=as.numeric(linear_trend_model$coefficients[2])  

  ggplot(df_energy_data, aes(x=date, y=df_energy_data[,3])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(df_energy_data)[3],sep="")) +
            geom_smooth(color="red",method="lm")
##Hydroelectric
  linear_trend_model=lm(df_energy_data[,4]~t)  
  summary(linear_trend_model)
  Hydrobeta0=as.numeric(linear_trend_model$coefficients[1])  
  Hydrobeta1=as.numeric(linear_trend_model$coefficients[2])  

  ggplot(df_energy_data, aes(x=date, y=df_energy_data[,4])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(df_energy_data)[4],sep="")) +
            geom_smooth(color="red",method="lm")
```
##Both total renewable energy production and total biomass energy production present upward trends, while total hydroeoectric power shows a downward trend.

### Q3
Use the *lm()* function to fit a linear trend to the three time series. Ask R to print the summary of the regression. Interpret the regression output, i.e., slope and intercept. Save the regression coefficients for further analysis.

```{r Q3}
Biolm=lm(ts_energy_data[,1]~t)
summary(Biolm)
Renewlm=lm(ts_energy_data[,2]~t)
summary(Renewlm)
Hydrolm=lm(ts_energy_data[,3]~t)
summary(Hydrolm)

```
##The linear trend for the time series of biomass energy production has an intercept of 37.44854, and a slope of 0.86734. At the beginning of the series, the value is 37.44854 units. After each time period, the production of biomass energy is expected to increase by 0.86734 units. Both the intercept and the slope have small p-values, meaning that they are statistically significant.
##For renewable energy production, the intercept is 98.07627 and the slope is 0.66527, meaning the start of the series has a value of 98.07627 units, and after each period the value is expected to increase 0.66527 units. Both of the intercept and the slope are statistically significant.
##While for hydroelectric energy consumption, the starting value is 385.3595 units, and after each period, it is expected to decrease by 0.3152 units. Both of the intercept and the slope are statistically significant.

### Q4
Use the regression coefficients from Q3 to detrend the series. Plot the detrended series and compare with the plots from Q1. What happened? Did anything change?

```{r}
##Biomass
detrend_Bioenergy_data <- df_energy_data[,2]-(Biobeta0+Biobeta1*t)
  par(mfrow=c(1,3))
  ggplot(df_energy_data, aes(x=date, y=df_energy_data[,2])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(df_energy_data)[2],sep="")) +
            geom_smooth(color="red",method="lm") +
            geom_line(aes(y=detrend_Bioenergy_data), col="green")+
            geom_smooth(aes(y=detrend_Bioenergy_data),color="orange",method="lm")
##Renewable
detrend_Renewenergy_data <- df_energy_data[,3]-(Renewbeta0+Renewbeta1*t)
  par(mfrow=c(1,3))
  ggplot(df_energy_data, aes(x=date, y=df_energy_data[,3])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(df_energy_data)[3],sep="")) +
            geom_smooth(color="red",method="lm") +
            geom_line(aes(y=detrend_Renewenergy_data), col="green")+
            geom_smooth(aes(y=detrend_Renewenergy_data),color="orange",method="lm")
##Hydroelectric
detrend_Hydroenergy_data <- df_energy_data[,4]-(Hydrobeta0+Hydrobeta1*t)
  par(mfrow=c(1,3))
  ggplot(df_energy_data, aes(x=date, y=df_energy_data[,4])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(df_energy_data)[4],sep="")) +
            geom_smooth(color="red",method="lm") +
            geom_line(aes(y=detrend_Hydroenergy_data), col="green")+
            geom_smooth(aes(y=detrend_Hydroenergy_data),color="orange",method="lm")
```
##The detrended data shows a regression line with a 0 value slope. It removes the overall increase or decrease of the time series and shows different patterns.

### Q5
Plot ACF and PACF for the detrended series and compare with the plots from Q1. Did the plots change? How?

```{r Q5}
    par(mfrow=c(1,3))
  plot(detrend_Bioenergy_data,type="l",col="black",ylab="Production [Trillion Btu]",main="Biomass Energy Production")
  Acf(detrend_Bioenergy_data,lag.max=40,main=paste("Energy Production",i,sep="")) 
  Pacf(detrend_Bioenergy_data,lag.max=40,main=paste("Energy Production",i,sep=""))
  
    par(mfrow=c(1,3))
  plot(detrend_Renewenergy_data,type="l",col="black",ylab="Production [Trillion Btu]",main="Renewable Energy Production")
  Acf(detrend_Renewenergy_data,lag.max=40,main=paste("Energy Production",i,sep="")) 
  Pacf(detrend_Renewenergy_data,lag.max=40,main=paste("Energy Production",i,sep=""))
  
    par(mfrow=c(1,3))
  plot(detrend_Hydroenergy_data,type="l",col="black",ylab="Production [Trillion Btu]",main="Hydroelectric Energy Consumption")
  Acf(detrend_Hydroenergy_data,lag.max=40,main=paste("Energy Consumption",i,sep="")) 
  Pacf(detrend_Hydroenergy_data,lag.max=40,main=paste("Energy Consumption",i,sep=""))

```
##Both the acf and pacf graphs for the three series showed smaller absolute values comparing to the original graphs. Among the three series, Biomass and renewable energy production’s graphs have relatively greater degrees of decrease while the changes in hydroelectric’s acf and pacf graphs are not significant.

## Seasonal Component

Set aside the detrended series and consider the original series again from Q1 to answer Q6 to Q8.

### Q6
Do the series seem to have a seasonal trend? Which serie/series? Use function *lm()* to fit a seasonal means model (i.e. using the seasonal dummies) to this/these time series. Ask R to print the summary of the regression. Interpret the regression output. Save the regression coefficients for further analysis.

```{r Q6Bio}

Biodummies <- seasonaldummy(ts_energy_data[,1])  

Bioseas_means_model=lm(df_energy_data[,2]~Biodummies)
summary(Bioseas_means_model)

Biobeta_int=Bioseas_means_model$coefficients[1]
Biobeta_coeff=Bioseas_means_model$coefficients[2:12]

Bioenergy_seas_comp=array(0,nobservation)
for(i in 1:nobservation){
  Bioenergy_seas_comp[i]=(Biobeta_int+Biobeta_coeff%*%Biodummies[i,])
}

ggplot(df_energy_data, aes(x=date, y=df_energy_data[,2])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(df_energy_data)[2],sep="")) +
            geom_line(aes(y=Bioenergy_seas_comp), col="red")

```
```{r Q6Renew}

Renewdummies <- seasonaldummy(ts_energy_data[,2])  

Renewseas_means_model=lm(df_energy_data[,3]~Renewdummies)
summary(Renewseas_means_model)

Renewbeta_int=Renewseas_means_model$coefficients[1]
Renewbeta_coeff=Renewseas_means_model$coefficients[2:12]

Renewenergy_seas_comp=array(0,nobservation)
for(i in 1:nobservation){
  Renewenergy_seas_comp[i]=(Renewbeta_int+Renewbeta_coeff%*%Renewdummies[i,])
}

ggplot(df_energy_data, aes(x=date, y=df_energy_data[,3])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(df_energy_data)[3],sep="")) +
            geom_line(aes(y=Renewenergy_seas_comp), col="red")

```

```{r Q6Hydro}

Hydrodummies <- seasonaldummy(ts_energy_data[,3])  

Hydroseas_means_model=lm(df_energy_data[,4]~Hydrodummies)
summary(Hydroseas_means_model)

Hydrobeta_int=Hydroseas_means_model$coefficients[1]
Hydrobeta_coeff=Hydroseas_means_model$coefficients[2:12]

Hydroenergy_seas_comp=array(0,nobservation)
for(i in 1:nobservation){
  Hydroenergy_seas_comp[i]=(Hydrobeta_int+Hydrobeta_coeff%*%Hydrodummies[i,])
}

ggplot(df_energy_data, aes(x=date, y=df_energy_data[,4])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(df_energy_data)[4],sep="")) +
            geom_line(aes(y=Hydroenergy_seas_comp), col="red")

```
##All three series show some seasonalities, with hydroelectric energy consumption being the only one that indicates strong seasonality: the regression outputs for biomass and renewable have R-squared values less than 0.05, while hydroelectric’s regression output has R-squared of 0.4182. The regression coefficients for hydroelectric energy indicate there’s above December consumption from March to July and January, and below December consumption from August to November and February.

### Q7
Use the regression coefficients from Q6 to deseason the series. Plot the deseason series and compare with the plots from part Q1. Did anything change?

```{r Q7}

Biodeseason_energy_data <- df_energy_data[,2]-Bioenergy_seas_comp

  ggplot(df_energy_data, aes(x=date, y=df_energy_data[,2])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(df_energy_data)[2],sep="")) +
            geom_line(aes(y=Biodeseason_energy_data), col="green")

Renewdeseason_energy_data <- df_energy_data[,3]-Renewenergy_seas_comp

  ggplot(df_energy_data, aes(x=date, y=df_energy_data[,3])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(df_energy_data)[3],sep="")) +
            geom_line(aes(y=Renewdeseason_energy_data), col="green")

Hydrodeseason_energy_data <- df_energy_data[,4]-Hydroenergy_seas_comp

  ggplot(df_energy_data, aes(x=date, y=df_energy_data[,4])) +
            geom_line(color="blue") +
            ylab(paste0(colnames(df_energy_data)[4],sep="")) +
            geom_line(aes(y=Hydrodeseason_energy_data), col="green")
```
##There’s no significant changes in the graphs for biomass energy production and renewable energy production. For hydroelectric energy consumption, the graph displays less fluctuations between the two neighboring values than the original one.

### Q8
Plot ACF and PACF for the deseason series and compare with the plots from Q1. Did the plots change? How?

```{r Q8}
par(mfrow=c(1,3))
  plot(Biodeseason_energy_data,type="l",col="black",ylab="Production [Trillion Btu]",main="Biomass Energy Production")
  Acf(Biodeseason_energy_data,lag.max=40,main=paste("Energy Production",i,sep="")) 
  Pacf(Biodeseason_energy_data,lag.max=40,main=paste("Energy Production",i,sep=""))
  
par(mfrow=c(1,3))
  plot(Renewdeseason_energy_data,type="l",col="black",ylab="Production [Trillion Btu]",main="Renewable Energy Production")
  Acf(Renewdeseason_energy_data,lag.max=40,main=paste("Energy Production",i,sep="")) 
  Pacf(Renewdeseason_energy_data,lag.max=40,main=paste("Energy Production",i,sep=""))
  
par(mfrow=c(1,3))
  plot(Hydrodeseason_energy_data,type="l",col="black",ylab="Production [Trillion Btu]",main="Hydroelectric Energy Consumption")
  Acf(Hydrodeseason_energy_data,lag.max=40,main=paste("Energy Consumption",i,sep="")) 
  Pacf(Hydrodeseason_energy_data,lag.max=40,main=paste("Energy Consumption",i,sep=""))

```
##The acf graphs shows much smoother decline than the original graphs. While not much changed for the biomass and renewable graphs, the acf graph for hydroelectric eliminates the characteristic of seasonality, as the values no longer fluctuates around the zero line. The pacf graph for hydroelectric also displays smaller absolute values, implying weaker partial correlation between the series and the lags itself.


